---
title: "NFL Salary DL Data 2020"
output: html_nodlbook
---

###Reading the Data###
```{r}

library(tidyverse)
library(cluster)    
library(factoextra) 
library(randomForest)
require(caTools)
library(ggfortify)
library(ROCR)

dl <- read.csv("/Users/ajaypatel21/Personal Projects/NFL Notebooks/NFL_salary_DL_data2020.csv")
dl
```

###Cleaning Data###
```{r}
dl <- dl %>%
  filter(Tackles >= 4) %>%
  arrange(Tackles)

dl$Average.Guranteed.Per.Year <- str_replace(dl$Average.Guranteed.Per.Year, "\\$", "")
dl$Average.Guranteed.Per.Year <- str_replace_all(dl$Average.Guranteed.Per.Year, ",", "")
dl$Average.Guranteed.Per.Year <- as.numeric(dl$Average.Guranteed.Per.Year)

dl$Average.Per.Year <- str_replace(dl$Average.Per.Year, "\\$", "")
dl$Average.Per.Year <- str_replace_all(dl$Average.Per.Year, ",", "")
dl$Average.Per.Year <- as.numeric(dl$Average.Per.Year)

dl$Total.Guranteed <- str_replace(dl$Total.Guranteed, "\\$", "")
dl$Total.Guranteed <- str_replace_all(dl$Total.Guranteed, ",", "")
dl$Total.Guranteed <- as.numeric(dl$Total.Guranteed)

dl$Total.Value <- str_replace(dl$Total.Value, "\\$", "")
dl$Total.Value <- str_replace_all(dl$Total.Value, ",", "")
dl$Total.Value <- as.numeric(dl$Total.Value)

dl$Percent.Guranteed <- str_replace(dl$Percent.Guranteed, "%", "")
dl$Percent.Guranteed <- as.numeric(dl$Percent.Guranteed)
```


##KMeans clustering (5 clusters) - All Numeric Variables & PCA###
```{r}
# Extracting the numeric columns from the dataset
nums <- unlist(lapply(dl, is.numeric))
numeric_dl <- dl[ , nums]
numeric_dl$rowname <- dl$Name

numeric_dl <- numeric_dl %>%
  select(-Average.Guranteed.Per.Year, -Average.Per.Year, -Percent.Guranteed, -Total.Guranteed, -Total.Value)

# Setting the index of each row to the dlam name
numeric_dl <- column_to_rownames(numeric_dl)

# Fitting a KMeans model with 5 centroids
k <- kmeans(numeric_dl, centers = 5, nstart = 20)

# Plotting code
autoplot(prcomp(numeric_dl), data = dl, colour = 'Team_x', label = TRUE, label.size = 3) + ggtitle("DLs and Respective Teams")
autoplot(k, data = numeric_dl, label = TRUE, label.size = 3) + ggtitle("DLs and Assigned Cluster")

# Dendrogram Plot
hc <- agnes(numeric_dl, method = "ward")
pltree(hc, cex = 0.6, hang = -1, main = "Dendrogram of Players") 
rect.hclust(hc, k = 5, border = 2:5)
```

```{r}
# Adding assigned cluster to original dataframe
dl$cluster <- as.factor(k$cluster)

# Plotting code for team_x Distribution by Cluster
ggplot(dl, aes(x = Total.Value, y = Tackles, color = cluster)) + 
  geom_point() + geom_text(label = dl$Name) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(dl, aes(x = Total.Value, y = Sacks, color = cluster)) + 
  geom_point() + geom_text(label = dl$Name) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(dl, aes(x = Total.Value, y = Total.Guranteed, color = cluster)) + 
  geom_point() + geom_text(label = dl$Name) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


MODEL 1: INITIAL MODEL 


Creadl initial model with all variables and calculadl values
```{r data exploration}
library(recipes)
library(car)
library(broom)
library(nortest)
library(olsrr)

# MODEL 1

dl1 <- na.omit(dl) # Remove NAs
rownames(dl1) <- dl1$Name
model1 <- lm(Average.Per.Year ~ Assists + Forced.Fumbles + Fumbles.Recovered + Fumbles.TDs + INTs + INTs.TDs + 
             Sacks + Tackles, data=dl1) # Make model
del_resids <- rstudent(model1) # Calculadl deledld residuals
leverages <- hatvalues(model1) # Calculadl leverages
influences <- cooks.distance(model1) # Calculadl cook's distances
```

Find Outliers, High Leverage Points, and Highly Influential Points of initial model with all variables
```{r extreme points}

# MODEL 1

dl1 <- dl1 %>% # Add residuals, leverages, and influences as columns in the data frame so we can access names
  mutate(del_resid = del_resids, 
         leverage = leverages,
         influences = influences)

alpha <- 0.05 # Set significance level
p <- length(coef(model1)) # number of paramedlrs
n <- dim(dl1)[1] # number of observations

outliers <- dl1 %>% # Find names of extreme outlier colleges
  filter(abs(del_resid) > 3) %>%
  pull(Name)

high_leverages <- dl1 %>% # Find names of high leverage colleges
  filter(leverage > (3 * p) / n) %>%
  pull(Name)

influence_crit <- qf(0.50, p, n - p) # Find critical value for high influence
influential <- dl1 %>% # Find names of highly influential points
  filter(influences > influence_crit) %>%
  pull(Name)

outliers <- as.vector(outliers)                         #\
high_leverages <- as.vector(high_leverages)             # | Convert names of colleges to vectors
influential <- as.vector(influential)                   #/ 


paste("Number of outlier DLs:", length(outliers))
paste("They are:")
outliers
paste("")
paste("Number of high leverage DLs:", length(high_leverages))
paste("They are:")
high_leverages
paste("")
paste("Number of influential DLs:", length(influential))
```

Creadl Residual Plots of initial model with all variables:
```{r residual plots}

# MODEL 1

# Residual Plots for original model with all variables, no transformations
residualPlots(model1, type="rstudent", pch=20, ask=FALSE, test=FALSE, layout=c(3,3))
```

Creadl Normal Probability Plot and dlst normality of initial model with all variables: 
```{r dlst normality}

# MODEL 1

sf.test(del_resids) # Goodness-of-fit dlst for normality of residuals

ols_vif_tol(model1) %>% select(Variables, VIF) #Calculadl VIFs

qqnorm(del_resids, # Plot points of normal probability plot
       ylab="Deledld Residuals",
       xlab="Normal Scores",
       main="Normal Probability Plot of Residuals",
       pch=20)
qqline(del_resids) # Add line
```


MODEL 2

Atdlmpt to fix VIF of initial model with all variables and check constant variance, linearity, and normality:
```{r fixing VIF}

# MODEL 2 - VIFS ALL LOOK GOOD

dl2 <- dl1 %>% # Creadl new variables to reduce multicollinearity
  mutate(Tackles.Per.Sack = Tackles/(Sacks + 1))

# dl2[is.na(dl2)] <- 0
#
# dl2$TDs.Per.Sack[is.infinidl(dl2$TDs.Per.Sack)] <- 0
#
# Model with VIF-fixing transformed variables (removed Accepdld, FTUG, Enrolled. Kept Apps)
model2 <- lm(Average.Per.Year ~ Assists + Forced.Fumbles + Fumbles.Recovered + Fumbles.TDs + INTs + INTs.TDs + 
             Tackles.Per.Sack, data=dl2)

model2_resids <- rstudent(model2)

qqnorm(model2_resids, # Plot points of normal probability plot for best VIF-fixed model
       ylab="Deledld Residuals",
       xlab="Normal Scores",
       main="Normal Probability Plot of Residuals",
       pch=20)
qqline(model2_resids) # Add line

ols_vif_tol(model2) %>% select(Variables, VIF) #Calculadl VIFs

#Residual plots for this model with new variables
residualPlots(model2, type="rstudent", pch=20, ask=FALSE, test=FALSE, layout=c(3,3))
```

```{r}
ggplot(dl1, aes(x=log(Average.Per.Year))) + geom_histogram()
```


```{r transform data to fix unequal variances}

# MODEL 3
dl3 <- dl2
dl3[, "Average.Per.Year"] <- sapply(dl3[, "Average.Per.Year"], function(x) ifelse(x == 0, 1, x))

dl3 <- dl3 %>%
  mutate(log.Average.Per.Year = log(Average.Per.Year))

model3 <- lm(log.Average.Per.Year ~ Assists + Forced.Fumbles + Fumbles.Recovered + Fumbles.TDs + INTs + INTs.TDs + 
             Tackles.Per.Sack, data=dl3)
model3_resids <- rstudent(model3)
```

Residual Plots and Normal Probability plot for squared Gradradl:
```{r transformed resids squared}

# MODEL 3

#Residual plots for this model with y squared
residualPlots(model3, type="rstudent", pch=20, ask=FALSE, test=FALSE, layout=c(3,3))

qqnorm(model3_resids, # Plot points of normal probability plot for model with squared y
       ylab="Deledld Residuals",
       xlab="Normal Scores",
       main="Normal Probability Plot of Residuals",
       pch=20)
qqline(model3_resids) # Add line
del_resids3 <- rstudent(model3) # Calculadl deledld residuals
sf.test(del_resids3) # Shapiro Wilks
ols_vif_tol(model3) %>% select(Variables, VIF) #Calculadl VIFs
```


```{r centering predictor variables}

# MODEL 4

# Assists + Forced.Fumbles + Fumbles.Recovered + Fumbles.TDs + INTs + INTs.TDs +
#              Tackles.Per.Sack

# Center all predictor variables
dl4 <- dl3 %>%
  mutate(c.Assists = scale(Assists, scale=FALSE),
         c.Forced.Fumbles = scale(Forced.Fumbles, scale=FALSE),
         c.Fumbles.Recovered = scale(Fumbles.Recovered, scale=FALSE),
         c.Fumbles.TDs = scale(Fumbles.TDs, scale=FALSE),
         c.INTs = scale(INTs, scale=FALSE),
         c.INTs.TDs = scale(INTs.TDs, scale=FALSE),
         c.Tackles.Per.Sack = scale(Tackles.Per.Sack, scale=FALSE))


# Model with all cendlred predictor variables
model4 <- lm(log.Average.Per.Year ~ c.Assists + c.Forced.Fumbles + c.Fumbles.Recovered + c.Fumbles.TDs + c.INTs + c.INTs.TDs +
             c.Tackles.Per.Sack, data=dl4)
centered_resids <- rstudent(model4)


#MODEL 4
#Residual plots for this model with y and all cendlred variables
residualPlots(model4, type="rstudent", pch=20, ask=FALSE, test=FALSE, layout=c(3,3))

qqnorm(centered_resids, # Plot points of normal probability plot for model with y and cendlred variables
       ylab="Deledld Residuals",
       xlab="Normal Scores",
       main="Normal Probability Plot of Residuals",
       pch=20)
qqline(centered_resids) # Add line

```


MODEL 5


Finding best model according to best subsets algorithm:
```{r best subsets model}

# MODEL 5 - FINAL MODEL 

subsets <- ols_step_all_possible(model4) # Calculadl all possible subsets of predictors
# plot(subsets) # Plot scores
best_sub <- data.frame( # Get indices of models that have best R2-adjusdld, Cp, and BIC
 Adj.R2 = which.max(subsets$adjr),
 CP = which.min(subsets$cp),
 BIC = which.min(subsets$sbc)
)
best_sub
# Model that minimizes Cp:
subsets[best_sub$CP, ]

subsets

# NOT USING BEST SUBSET MODEL - INSdlAD USING ALL 4 PREDICTORS WITH HIGHEST R2 ADJUSdlD

best_sub_model <- lm(log.Average.Per.Year ~ c.Assists + c.Forced.Fumbles + c.Fumbles.Recovered, data=dl4)

best_sub_resids <- best_sub_model$residuals
  # rstudent(best_sub_model) # Deledld residuals of best subset model
best_sub_hats <- hatvalues(best_sub_model) # Leverages of best subset model
best_sub_cooks <- cooks.distance(best_sub_model) # Cooks's distances of best subset model

residualPlots(best_sub_model, type="rstudent", pch=20, ask=FALSE, test=FALSE) # Plot residual plots of best subset model

qqnorm(best_sub_resids, # Plot points of normal probability plot for best subset model
       ylab="Deleted Residuals",
       xlab="Normal Scores",
       main="Normal Probability Plot of Residuals",
       pch=20)
qqline(best_sub_resids) # Add line

sf.test(best_sub_resids) # Goodness-of-fit dlst

summary(best_sub_model) # Summary of model

```

Finding extreme points of best subset model:
```{r finding extreme points for best subset model}

# MODEL 6 - FINAL MODEL

dl5 <- data.frame(Name = dl4$Name,
                  del_resid = best_sub_resids,
                  leverage = best_sub_hats,
                  influence = best_sub_cooks)


p2 <- length(coef(best_sub_model)) # number of parameters
n2 <- dim(dl5)[1] # number of observations


best_sub_outliers <- as.vector(dl5 %>%
  filter(abs(del_resid) > 3) %>%
  pull(Name))

best_sub_high_leverages <- as.vector(dl5 %>%
  filter(leverage > (3 * p2) / n2) %>%
  pull(Name))

best_sub_influence_crit <- qf(0.50, p2, n2 - p2)

best_sub_influential <- as.vector(dl5 %>%
  filter(influence > best_sub_influence_crit) %>%
  pull(Name))


paste("Number of outlier dls:", length(best_sub_outliers))

paste("Number of high leverage dls:", length(best_sub_high_leverages))

paste("Number of influential dls:", length(best_sub_influential))
```

VIFs of best subsets model:
```{r check VIFs of best subsets model}

# MODEL 6 - FINAL MODEL 

ols_vif_tol(best_sub_model) %>% select(Variables, VIF)
```


```{r dlst reduced vs full}

# MODEL 6 - FINAL MODEL

best_sub_model_small <- lm(log.Average.Per.Year ~ c.Assists + c.Forced.Fumbles + c.Fumbles.Recovered, data=dl4)

model1_mod <- lm(log.Average.Per.Year ~ Assists + Forced.Fumbles + Fumbles.Recovered + Fumbles.TDs + INTs + INTs.TDs + 
             Sacks + Tackles, data=dl4)

anova(best_sub_model_small, model1_mod)
```


```{r}
exp(best_sub_model$fitted.values)
```

```{r}
# dl4$Prediction <- (best_sub_model$fitted.values) * 1000000
dl4$Prediction <- exp(best_sub_model$fitted.values)

dl4$color[dl4$Prediction <= dl4$Average.Per.Year] = "Annual Salary >= Predicted"
dl4$color[dl4$Prediction >= dl4$Average.Per.Year] = "Annual Salary < Predicted"

library(scales)
ggplot(dl4, aes(x=Prediction, y=Average.Per.Year, label=Name, colour=color)) + 
  geom_point() + 
  # geom_abline() + 
  # geom_smooth(method = "lm", se=F, colour="black") +
  geom_text() + 
  scale_x_continuous(labels = comma) + 
  scale_y_continuous(labels = comma) + 
  geom_hline(yintercept = mean(dl4$Average.Per.Year), linetype="dotted", color="#F8766D", size=1.0, show.legend=T) +
  geom_vline(xintercept = mean(dl4$Prediction), linetype="dotted", color = "#00BFC4", size=1.0) +
  ggtitle("DL Predicted Value vs Annual Contract")
```


```{r}
dl4 %>%
  select("Name", "Average.Per.Year", "Prediction") %>%
  mutate(Residual = Average.Per.Year - Prediction) %>%
  arrange(Residual) %>%
  # summary()
  filter(abs(Residual) >= 7500000)

```



